Is it time for an NBA expansion?

### 2024-04-15 paper/paper.qmd
Created Model

library(tidyverse)
library(readr)

@@

```{r}
#| include: false
#| warning: false
#| message: false
#| 
#combining data into one table
combined_data <- data.frame()

@@

  # Combine with the main data frame
  combined_data <- bind_rows(combined_data, yearly_data)
}
#Remove Less important columns
full_nba_data <- combined_data %>% select(-"G", -"MP", -"2P", -"2PA", -"2P%", -"Rk")

#creating table for league average
nba_data <- full_nba_data %>% 
  filter(Team == "League Average")

#saving Clean Data
write.csv(nba_data, "../data/clean_data/nba_data.csv", row.names = FALSE)
write.csv(full_nba_data, "../data/clean_data/full_nba_data.csv", row.names = FALSE)
```

@@

>>>>

library(tidyverse)
library(readr)
library(modelsummary)

@@

combined_data <- data.frame()

@@

  # Combine with the main data frame
  combined_data <- bind_rows(combined_data, yearly_data)
}

# Add the number of teams to the combined data before selecting columns
combined_data <- combined_data %>%
  mutate(Num_Teams = case_when(
    Year <= 1980 ~ 22,
    Year > 1980 & Year <= 1988 ~ 23,
    Year == 1989 ~ 25,  
    Year > 1989 & Year <= 1995 ~ 27,
    Year > 1995 & Year <= 2004 ~ 29,
    TRUE ~ 30  # For years after 2004
  ))

# Now remove less important columns
full_nba_data <- combined_data %>%
  select(-G, -MP, -`2P`, -`2PA`, -`2P%`, -Rk)

# Creating table for league average
nba_data <- full_nba_data %>% 
  filter(Team == "League Average")

# Saving Clean Data
write.csv(nba_data, "../data/clean_data/nba_data.csv", row.names = FALSE)
write.csv(full_nba_data, "../data/clean_data/full_nba_data.csv", row.names = FALSE)
```

@@

```{r}
#| warning: false
#| message: false

# Simulating data
set.seed(123)  # For reproducibility
sim_nba_data <- data.frame(
  Year = rep(1980:2023, times = 10),  # Simulate 10 records per year
  Num_teams = c(rep(22, each = 1 * 10),
                rep(23, each = 8 * 10),
                rep(25, each = 1 * 10),
                rep(27, each = 6 * 10),
                rep(29, each = 9 * 10),
                rep(30, each = 19 * 10)),  # Adjusted each parameter as needed
  PTS = rnorm(440, mean = mean(full_nba_data$PTS), sd = sd(full_nba_data$PTS)),  
  AST = rnorm(440, mean = mean(full_nba_data$AST), sd = sd(full_nba_data$AST)), 
  TRB = rnorm(440, mean = mean(full_nba_data$TRB), sd = sd(full_nba_data$TRB)),  
  STL = rnorm(440, mean = mean(full_nba_data$STL), sd = sd(full_nba_data$STL)),  
  BLK = rnorm(440, mean = mean(full_nba_data$BLK), sd = sd(full_nba_data$BLK)),  
  TOV = rnorm(440, mean = mean(full_nba_data$TOV), sd = sd(full_nba_data$TOV)) 
)
# Add a column for PPG that depends on the year and number of teams, 
# as well as other factors. Adjust the coefficients as needed for your simulation.
sim_nba_data_1 <- sim_nba_data %>%
  mutate(PPG = (PTS / 82) + (Year - 1980) * 0.2 - (Num_teams - 22) * 0.1 +
                0.1 * AST + 0.05 * TRB - 0.05 * STL + 0.05 * BLK - 0.05 * TOV)

# Fit a multiple linear regression model
nba_model <- lm(PTS ~ Year * Num_teams + AST + TRB + STL + BLK + TOV, data = sim_nba_data)
nba_model_1 <- lm(PTS ~ Year * Num_teams + AST + TRB + STL + BLK + TOV, data = sim_nba_data_1)

# Summary of the linear model
msummary(nba_model)
modelsummary(nba_model_1)
